Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-04-25 doc/hiv-dynamics/presentation.tex

\documentclass{beamer}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{gillius}
\usepackage{abraces}
\usepackage{etex}
\usepackage{array}
\usepackage{multirow}
\usepackage{qrcode}
\usepackage[backend=biber]{biblatex}

@@

\begin{frame}{Most epidemiological models assume homogeneous mixing}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \centering
      \includegraphics{compartments}
      \vspace{1cm}

      \uncover<2->{\includegraphics[width=\textwidth]{sir-trajectories}}
    \end{column}
    \begin{column}{0.4\textwidth}
      \vspace{-2cm}
      \begin{itemize}
        \setlength{\itemsep}{24pt}
        \item Often provides reasonable approximation in practice.
        \uncover<2->{
        \item Can be inaccurate when substantial contact heterogeneity exists.
        }
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}{Scale-free networks and preferential attachment}
  \includegraphics[width=\textwidth]{pa-example}
  \begin{itemize}
    \item Sexual networks are often scale-free: degree distributions appear
      follow a power law ($\Pr(\text{degree} = d) \propto d^{\gamma}$).
      \pause
    \item Preferential attachment models generate scale-free networks.
  \end{itemize}
\end{frame}

\begin{frame}{Barab\'asi-Albert (BA) network model}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{minipage}[p][\textheight][t]{\textwidth}
        \centering
        \only<1>{\includegraphics[page=1, width=\textwidth]{pa.pdf}}
        \only<2>{\includegraphics[page=2, width=\textwidth]{pa.pdf}}
        \only<3>{\includegraphics[page=3, width=\textwidth]{pa.pdf}}
        \only<4>{\includegraphics[page=4, width=\textwidth]{pa.pdf}}
        \only<5>{\includegraphics[page=5, width=\textwidth]{pa.pdf}}
        \only<6>{\includegraphics[page=6, width=\textwidth]{pa.pdf}}
      \end{minipage}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{minipage}[t][\textheight][t]{\textwidth}
      \begin{itemize}
        \uncover<1->{ \item Start with a small number of connected nodes. }
        \uncover<2->{ \item Attach new nodes with $m$ edges. }
        \uncover<3->{ \item Other endpoints of degree $d$ are chosen with probability $\propto d^{\alpha} + 1$. }
        \uncover<5->{ \item Continue until network has $N$ nodes. }
        \uncover<6->{ \item Also consider the prevalence $I$ when the transmission
          tree is sampled.}
      \end{itemize}
      \end{minipage}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Do network parameters measurably affect tree shape?}
  \only<1>{
    \includegraphics[height=1in, trim=0 2in 4in 0, clip]{kernel-idea}
  }
  \only<2>{
    \includegraphics[height=1in, trim=0 2in 2in 0, clip]{kernel-idea}
  }
  \only<3->{
    \includegraphics[height=1in, trim=0 2in 0 0, clip]{kernel-idea}
  }
  \begin{itemize}
    \item Generate networks under different parameter values ($N$, $m$, $\alpha$).
    \uncover<2->{
    \item Simulate epidemic over each network with Gillespie algorithm ($I$).
    }
    \uncover<3->{
    \item Randomly subsample to form transmission trees.
    }
    \uncover<4->{
    \item Compare pairwise using tree kernel\footfullcite{poon2013mapping}.
    }
  \end{itemize}
\end{frame}

\begin{frame}{$\alpha$ and $I$ separate well in feature space}
  \begin{tikzpicture}
    \node {\includegraphics[height=0.8\textheight]{kernel-kpca.pdf}};
    \fill<1>[white] (-4.6,-3.2) rectangle (5,0.2);
  \end{tikzpicture}
  \pause
\end{frame}

\begin{frame}{Classifiers for network parameters}
  \vspace{-1cm}
  \begin{itemize}
    \item kSVM using tree kernel
    \item SVM using normalized
      lineages-through-time\footfullcite{janzen2015approximate}
    \item linear regression using Sackin's index~\footfullcite{shao1990tree}
  \end{itemize}

  \pause
  \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
\end{frame}

\begin{frame}{Approximate Bayesian computation (ABC) for networks}
  \begin{minipage}[p][\textheight][t]{\textwidth}
    \only<1>{\includegraphics[width=\textwidth, trim=0 3in 0 0, clip]{abc-idea.pdf}}
    \only<2>{\includegraphics[width=\textwidth, trim=0 1.8in 0 0, clip]{abc-idea.pdf}}
    \only<3>{\includegraphics[width=\textwidth, trim=0 1in 0 0, clip]{abc-idea.pdf}}
    \only<4>{\includegraphics[width=\textwidth, trim=0 0in 0 0, clip]{abc-idea.pdf}}
  \end{minipage}
\end{frame}

\begin{frame}{Marginal posterior distributions for one simulated tree}
  \begin{minipage}[p][\textheight][t]{\textwidth}
    \only<1>{\centerline{\includegraphics[width=0.9\textwidth, trim=0 2.5in 0 0, clip]{abc-posterior-example.pdf}}}
    \only<2>{\centerline{\includegraphics[width=0.9\textwidth]{abc-posterior-example.pdf}}}
  \end{minipage}
\end{frame}

\begin{frame}{Accuracy of maximum \textit{a posteriori} estimates for $m = 2$}
  \begin{minipage}[p][\textheight][t]{\textwidth}
    \vspace{-0.5cm}
    \only<1>{\includegraphics[height=0.9\textheight, trim=0 0 3.4in 0, clip]{abc-point-estimate-m2.pdf}}
    \only<2>{\includegraphics[height=0.9\textheight]{abc-point-estimate-m2.pdf}}
  \end{minipage}
\end{frame}

\begin{frame}{Application to published HIV datasets}
  \uncover<2->{
  \includegraphics[width=\textwidth]{realdata-hpd-bc-horiz}
  }
  \vspace{0.5cm}

  \scriptsize
  \input{\tablepath/realdata-bc}
  \normalsize
\end{frame}

\begin{frame}{Conclusions}
  \begin{itemize}
    \item Kernel-ABC is a viable method for fitting contact network models to
      phylogenetic data.
      \pause
    \item Some parameters of the Barab\'asi-Albert preferential attachment data
      have a measurable impact on tree shape and can be re-estimated using
      kernel-ABC.
      \pause
    \item The networks underlying real epidemics may be heterogeneous,
      underscoring the importance of considering network structure in
      phylodynamic analyses.
      \pause
  \end{itemize}
  \centerline{\qrcode{github.com/rmcclosk/netabc} github.com/rmcclosk/netabc}
\end{frame}

@@

      \vfill
    \end{column}
    \begin{column}{0.4\textwidth}
      \centering

      \includegraphics[width=3cm]{logos/cihr}
      \vspace{0.5cm}

@@

      \includegraphics[width=1cm]{logos/ubc}
      \vspace{0.5cm}

      \includegraphics[width=3cm]{logos/btp}
    \end{column}
  \end{columns}
\end{frame}
\end{document}

>>>>

\documentclass{beamer}
\usepackage{tikz}
\usepackage{gillius}
\usepackage{abraces}
\usepackage{etex}
\usepackage{array}
\usepackage{multirow}
\usepackage{qrcode}
\usepackage[backend=biber, style=authoryear, uniquelist=minyear, uniquename=false]{biblatex}

@@

\begin{frame}{Most epidemiological models assume homogeneous mixing}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \centering
      \includegraphics[scale=0.6]{compartments}
      \vspace{1cm}

      \uncover<3->{\includegraphics[width=\textwidth]{sir-trajectories}}
    \end{column}
    \begin{column}{0.4\textwidth}
      \vspace{-2cm}
      \begin{itemize}
        \setlength{\itemsep}{24pt}
        \uncover<2->{
        \item Often provide a reasonable approximation in practice.
        }
        \uncover<3->{
        \item Can be inaccurate when substantial contact heterogeneity exists.
        }
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

@@

%\begin{frame}{Preferential attachment generates scale-free networks}
%  \includegraphics[width=\textwidth]{pa-example}
%  \begin{itemize}
%    \item Sexual networks are often scale-free: degree distributions appear
%      follow a power law ($\Pr(\text{degree} = d) \propto d^{-\gamma}$).
%      \pause
%    \item Preferential attachment models generate scale-free networks.
%  \end{itemize}
%\end{frame}

\begin{frame}{Barab\'asi-Albert model incorporates preferential attachment}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{minipage}[p][\textheight][t]{\textwidth}
        \centering
        \only<1>{\includegraphics[page=1, width=\textwidth]{pa.pdf}}
        \only<2>{\includegraphics[page=2, width=\textwidth]{pa.pdf}}
        \only<3>{\includegraphics[page=3, width=\textwidth]{pa.pdf}}
        \only<4>{\includegraphics[page=4, width=\textwidth]{pa.pdf}}
        \only<5>{\includegraphics[page=5, width=\textwidth]{pa.pdf}}
        \only<6>{\includegraphics[page=6, width=\textwidth]{pa.pdf}}
      \end{minipage}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{minipage}[t][\textheight][t]{\textwidth}
      \begin{itemize}
        \uncover<1->{ \item Start with a small number of connected nodes. }
        \uncover<2->{ \item Attach new nodes with $m$ edges. }
        \uncover<3->{ \item Other endpoints of degree $d$ are chosen with probability $\propto d^{\alpha} + 1$. }
        \uncover<5->{ \item Continue until network has $N$ nodes. }
        \uncover<6->{ \item Also consider the prevalence $I$. }
      \end{itemize}
      \end{minipage}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Do network parameters measurably affect tree shape?}
  \only<1>{
    \includegraphics[height=1in, trim=0 2in 4in 0, clip]{kernel-idea}
  }
  \only<2>{
    \includegraphics[height=1in, trim=0 2in 2in 0, clip]{kernel-idea}
  }
  \only<3->{
    \includegraphics[height=1in, trim=0 2in 0 0, clip]{kernel-idea}
  }
  \begin{itemize}
    \setlength{\itemsep}{0pt}
    \item Generate networks under different parameter values (number of nodes
      $N$, number of edges per vertex $m$, pereferential attachment power
      $\alpha$).
    \uncover<2->{
    \item Simulate epidemic over each network until $I$ nodes are infected.
    }
    \uncover<3->{
    \item Randomly subsample to form transmission trees.
    }
    \uncover<4->{
    \item Compare trees pairwise using tree kernel\footfullcite{poon2013mapping}.
    }
  \end{itemize}
\end{frame}

\begin{frame}{Preferential attachment power $\alpha$ affects tree shape}
  \begin{tikzpicture}[remember picture, overlay]
    \node at (current page.south west) [anchor=south west] {\includegraphics[width=0.6\textwidth, trim=0.2in 2.7in 3.3in 0, clip]{kernel-kpca.pdf}};
    \node (t) at (current page.north east) [anchor=north east, inner ysep=1.5cm, inner xsep=0.5cm] {\includegraphics[width=0.7\textwidth]{kernel-alpha-tree.pdf}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Prevalence $I$ affects tree shape}
  \begin{tikzpicture}[remember picture, overlay]
    \node at (current page.south west) [anchor=south west] {\includegraphics[width=0.6\textwidth, trim=3.5in 2.7in 0in 0, clip]{kernel-kpca.pdf}};
    \node (t) at (current page.north east) [anchor=north east, inner ysep=1.5cm, inner xsep=0.5cm] {\includegraphics[width=0.7\textwidth]{kernel-I-tree.pdf}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Number of edges per vertex $m$ does not affect tree shape}
  \begin{tikzpicture}[remember picture, overlay]
    \node at (current page.south west) [anchor=south west] {\includegraphics[width=0.6\textwidth, trim=0.2in 0.3in 3.3in 2.2in, clip]{kernel-kpca.pdf}};
    \node (t) at (current page.north east) [anchor=north east, inner ysep=1.5cm, inner xsep=0.5cm] {\includegraphics[width=0.7\textwidth]{kernel-m-tree.pdf}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Number nodes $N$ modestly affects tree shape}
  \begin{tikzpicture}[remember picture, overlay]
    \node at (current page.south west) [anchor=south west] {\includegraphics[width=0.6\textwidth, trim=3.5in 0.3in 0 2.2in, clip]{kernel-kpca.pdf}};
    \node (t) at (current page.north east) [anchor=north east, inner ysep=1.5cm, inner xsep=0.5cm] {\includegraphics[width=0.7\textwidth]{kernel-N-tree.pdf}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Simulated trees can be used to fit a network model}
  \begin{minipage}[p][\textheight][t]{\textwidth}
    \only<1>{\includegraphics[width=0.9\textwidth, trim=0 3in 0 0, clip]{abc-idea.pdf}}
    \only<2>{\includegraphics[width=0.9\textwidth, trim=0 1.8in 0 0, clip]{abc-idea.pdf}}
    \only<3>{\includegraphics[width=0.9\textwidth, trim=0 1in 0 0, clip]{abc-idea.pdf}}
    \uncover<4->{\includegraphics[width=0.9\textwidth, trim=0 0in 0 0, clip]{abc-idea.pdf}}
    \uncover<5->{\vspace{0.5cm}\centerline{\textbf{kernel approximate Bayesian computation (kernel-ABC)}}}
  \end{minipage}
\end{frame}

\begin{frame}{Example simulation shows $\alpha$ and $I$ can be reconstructed}
  \begin{minipage}[p][\textheight][t]{\textwidth}
    \begin{center}
    preferential attachment power \hfill prevalence
    \only<1>{
      \includegraphics[width=0.8\textwidth, trim=0 2.5in 0 0, clip]{abc-posterior-example.pdf}
    }
    \only<2>{
      \includegraphics[width=0.8\textwidth]{abc-posterior-example.pdf}

      \vspace{-0.25cm}
      edges per vertex \hfill total nodes
    }
    \end{center}
  \end{minipage}
\end{frame}

\begin{frame}{Maximum \textit{a posteriori} estimates are accurate for $\alpha$ and $I$}
  \begin{minipage}[p][\textheight][t]{\textwidth}
    \vspace{-0.5cm}
    preferential attachment power \hfill \uncover<2>{edges per vertex}

    \hspace{0.75cm}\only<1>{\includegraphics[height=0.8\textheight, trim=0 0 3.4in 0, clip]{abc-point-estimate-m2.pdf}}
    \only<2>{\includegraphics[height=0.8\textheight]{abc-point-estimate-m2.pdf}}

    prevalence \hfill \uncover<2>{total nodes}
  \end{minipage}
\end{frame}

\begin{frame}{Real world HIV datasets exhibit network heterogeneity}
  \uncover<2->{
    \centerline{\includegraphics[height=0.8\textheight]{realdata-hpd-bc}}
  }

%  \scriptsize
%    \begin{tabular}{ccccc}
%      Reference & Sequences ($n$) & Location & Risk group & Gene \\
%      \hline
%      \color{magenta} \textcite{novitsky2013phylogenetic} &
%        \color{magenta} \multirow{2}{*}{180} &
%        \color{magenta} \multirow{2}{*}{Mochudi, Botswana} & 
%        \color{magenta} \multirow{2}{*}{HET} &
%        \color{magenta} \multirow{2}{*}{\textit{env}} \\ 
%      \color{magenta} \textcite{novitsky2014impact} \\
%      \color{blue} \textcite{cuevas2009hiv} & 
%        \color{blue} 287 & 
%        \color{blue} Basque Country, Spain & 
%        \color{blue} mixed & 
%        \color{blue} \textit{pol} \\
%      \color{cyan} \textcite{li2015hiv} & 
%        \color{cyan} 280 & 
%        \color{cyan} Shanghai, China & 
%        \color{cyan} MSM & 
%        \color{cyan} \textit{pol} \\
%      \color{green} unpublished & 
%        \color{green} 399 & 
%        \color{green} British Columbia, Canada & 
%        \color{green} IDU & 
%        \color{green} \textit{pol} \\
%      \color{orange} \textcite{wang2015targeting} & 
%        \color{orange} 173 & 
%        \color{orange} Beijing, China & 
%        \color{orange} MSM & 
%        \color{orange} \textit{pol} \\
%      \color{red} \textcite{niculescu2015recent} & 
%        \color{red} 136 & 
%        \color{red} Romaina & 
%        \color{red} IDU & 
%        \color{red} \textit{pol} \\
%      \hline
%    \end{tabular}
%  \normalsize
\end{frame}

\begin{frame}{Conclusions}
  \begin{itemize}
    \item Kernel-ABC is the first phylodynamic method to fit contact network
      models to phylogenetic data.
      \pause
    \item The preferential attachment power of the Barab\'asi-Albert network
      model, which is challenging to estimate by traditional epidemiological
      methods, can be estimated with kernel-ABC.
      \pause
    \item The networks underlying real epidemics are heterogeneous,
      underscoring the importance of considering network structure in
      phylodynamic analyses.
      \pause
  \end{itemize}
  \centerline{\qrcode{github.com/rmcclosk/netabc} github.com/rmcclosk/netabc}
\end{frame}

@@

      \vfill
      \vspace{0.5cm}
      $\qquad\qquad$\includegraphics[width=2cm]{logos/genomecanada}
      $\qquad$
      \includegraphics[width=2cm]{logos/genomequebec}
    \end{column}
    \begin{column}{0.4\textwidth}
      \centering

      \includegraphics[width=3cm]{logos/cfe}
      \vspace{0.5cm}

      \includegraphics[width=3cm]{logos/cihr}

      \includegraphics[width=3cm]{logos/genomebc}

@@

      \includegraphics[width=3cm]{logos/btp}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Classifiers for network parameters}
  \vspace{-1cm}
  \begin{itemize}
    \item kSVR using tree kernel
    \item SVR using normalized
      lineages-through-time\footfullcite{janzen2015approximate}
    \item linear regression using Sackin's index~\footfullcite{shao1990tree}
  \end{itemize}

  \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
\end{frame}

\end{document}

### 2016-04-25 doc/tables/realdata.tex

\begin{tabular}{ccccc}
  Reference & Sequences ($n$) & Location & Risk group & Gene \\
  \hline
  \autocite{wang2015targeting} & 173 & Beijing, China & MSM & \textit{pol} \\
  \autocite{cuevas2009hiv} & 287 & Basque Country, Spain & mixed & \textit{pol} \\
  \autocite{novitsky2013phylogenetic} & \multirow{2}{*}{180} &
  \multirow{2}{*}{Mochudi, Botswana} & \multirow{2}{*}{mixed} &
  \multirow{2}{*}{\textit{env}} \\ \autocite{novitsky2014impact} \\
  \autocite{li2015hiv} & 280 & Shanghai, China & MSM & \textit{pol} \\
  \autocite{niculescu2015recent} & 136 & Romaina & IDU & \textit{pol} \\
  \hline
\end{tabular}

>>>>

\begin{tabular}{ccccc}
  Reference & Sequences ($n$) & Location & Risk group & Gene \\
  \hline
  \textcite{wang2015targeting} & 173 & Beijing, China & MSM & \textit{pol} \\
  \textcite{cuevas2009hiv} & 287 & Basque Country, Spain & mixed & \textit{pol} \\
  \textcite{novitsky2013phylogenetic} & \multirow{2}{*}{180} &
  \multirow{2}{*}{Mochudi, Botswana} & \multirow{2}{*}{HET} &
  \multirow{2}{*}{\textit{env}} \\ \textcite{novitsky2014impact} \\
  \textcite{li2015hiv} & 280 & Shanghai, China & MSM & \textit{pol} \\
  \textcite{niculescu2015recent} & 136 & Romania & IDU & \textit{pol} \\
  \hline
\end{tabular}

### 2016-04-25 doc/thesis/appendix.Rnw

>>>>

\begin{figure}
  \includegraphics{cuevas2009-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{cuevas2009hiv} data. Vertical lines indicate maximum \textit{a
      posteriori} estimates, and shaded areas are 95\% highest posterior
      density intervals. $x$-axis indicates regions of nonzero prior density.
  }
  \label{fig:cuevas}
\end{figure}

\begin{figure}
  \includegraphics{li2015-posterior}
  \caption{
      Posterior distribution of BA model parameters for \textcite{li2015hiv}
      data. Vertical lines indicate maximum \textit{a posteriori} estimates,
      and shaded areas are 95\% highest posterior density intervals. $x$-axis
      indicates regions of nonzero prior density.
  }
  \label{fig:li}
\end{figure}

\begin{figure}
  \includegraphics{niculescu2015-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{niculescu2015recent} data. Vertical lines indicate maximum
      \textit{a posteriori} estimates, and shaded areas are 95\% highest
      posterior density intervals. $x$-axis indicates regions of nonzero prior
      density.
  }
  \label{fig:niculescu}
\end{figure}

\begin{figure}
  \includegraphics{novitsky2014-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{novitsky2013phylogenetic, novitsky2014impact} data. Vertical
      lines indicate maximum \textit{a posteriori} estimates, and shaded areas
      are 95\% highest posterior density intervals. $x$-axis indicates regions
      of nonzero prior density.
  }
  \label{fig:novitsky}
\end{figure}

\begin{figure}
  \includegraphics{wang2015-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{wang2015targeting} data. Vertical lines indicate maximum
      \textit{a posteriori} estimates, and shaded areas are 95\% highest
      posterior density intervals. $x$-axis indicates regions of nonzero prior
      density.
  }
  \label{fig:wang}
\end{figure}

<<abc_posterior, echo=FALSE, results="asis">>=
    N <- 5000
    replicate <- 0
    for (m in c(2, 3, 4)) {
    for (I in c(1000, 2000)) {
    for (alpha in c(0, 0.5, 1, 1.5, 2)) {
        fn <- sprintf("abc-posterior/%2f_%d_%d_%d_%d.pdf", alpha, I, m, N, replicate)
        lab <- sprintf("fig:%2f-%d-%d-%d-%d", alpha, I, m, N, replicate)
        cat("\\begin{figure}\n")
        cat("\\centering\n")
        cat("\\includegraphics{", fn, "}\n")
        cat("\\caption{\n")
        cat("Approximate marginal posterior distributions of Barab\\'asi-Albert model parameters obtained using kernel-ABC for one simulated dataset. Dashed lines indicate true values, and shaded areas shown 95\\% highest posterior density intervals.\n")
        cat("}\n")
        cat("\\label{", lab, "}\n")
    }
    }
    }
@

### 2016-04-25 doc/thesis/methods.tex

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{gridsearch-expt}
  \caption{Graphical schematic of grid search experiments used to investigate 
    \gls{BA} model parameters.}
  \label{fig:gridexpt}
\end{figure}

@@

To test the full kernel-\gls{ABC} algorithm, we simulated four trees each under
a variety of parameter values, and ran the \software{netabc} program to
estimate posterior distributions for the parameters. The parameter values and
priors used are listed in \ref{tab:abcexpt}. The tree kernel meta-parameters
were set to $\lambda = 0.3$ and $\sigma = 4$. The \gls{SMC} algorithm was run
with 1000 particles, five sampled datasets per particle, and the $\alpha$
parameter (not to be confused with the \gls{BA} preferential attachment
parameter, see \cref{subsubsec:adaptsmc}) set to 0.95. The algorithm was
stopped when the acceptance rate of the \gls{MH} kernel dropped below 1.5\%,
the same criterion used by \citeauthor{del2012adaptive}. Approximate marginal
posterior densities for each parameter were calculated using the
\software{density} function in \software{R} applied to the final weighted
population of particles. Credible intervals were obtained for each parameter
using the \software{HPDinterval} function in the \software{coda}
package~\autocite{plummer2006coda}.

@@

We downloaded all sequences associated with each published study from GenBank.
For the \textcite{novitsky2014impact} data, each \textit{env} sequence was
aligned pairwise to the HXB2 reference sequence (GenBank accession number
K03455) and the hypervariable regions were clipped out with
\software{BioPython} version 1.66+~\autocite{cock2009biopython}. Sequences were
multiply aligned using \software{MUSCLE} version 3.8.31
\autocite{edgar2004muscle}, and alignments were manually inspected with
\software{Seaview} version 4.4.2 \autocite{gouy2010seaview}. Phylogenies were
constructed from the nucleotide alignments by approximate maximum likelihood
using \software{FastTree2} version 2.1.7 with the \gls{GTR} model. Transmission
trees were estimated by rooting and time-scaling the phylogenies by root-to-tip
regression, using a modified version of Path-O-Gen (distributed as part of
BEAST~\autocite{drummond2007beast}) as described
previously~\autocite{poon2015phylodynamic}.

>>>>

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{gridsearch-expt}
  \caption{
    Graphical schematic of grid search experiments used to investigate \gls{BA}
    model parameters. Trees were simulated along a narrowly spaced grid of
    parameter values (``grid trees''). Separate trees were simulated for a
    small subset of the grid values (``test trees''). Each test tree was
    compared to every grid tree using the tree kernel, and the resulting kernel
    scores were normalized to resemble a probability density from which the
    mode and 95\% highest density interval were calculated.
  }
  \label{fig:gridexpt}
\end{figure}

@@

To test the full kernel-\gls{ABC} algorithm, we simulated three trees each under
a variety of parameter values, and ran the \software{netabc} program to
estimate posterior distributions for the parameters. The parameter values and
priors used are listed in \cref{tab:abcexpt}. The tree kernel meta-parameters
were set to $\lambda = 0.3$ and $\sigma = 4$. The \gls{SMC} algorithm was run
with 1000 particles, five sampled datasets per particle, and the $\alpha$
parameter (not to be confused with the \gls{BA} preferential attachment
parameter, see \cref{subsubsec:adaptsmc}) set to 0.95. The algorithm was
stopped when the acceptance rate of the \gls{MH} kernel dropped below 1.5\%,
the same criterion used by \citeauthor{del2012adaptive}. Approximate marginal
posterior densities for each parameter were calculated using the
\software{density} function in \software{R} applied to the final weighted
population of particles. Credible intervals were obtained for each parameter
using the \software{HPDinterval} function in the \software{coda}
package~\autocite{plummer2006coda}.

@@

We downloaded all sequences associated with each published study from GenBank.
For the \textcite{novitsky2014impact} data, each \textit{env} sequence was
aligned pairwise to the HXB2 reference sequence (GenBank accession number
K03455) and the hypervariable regions were clipped out with
\software{BioPython} version 1.66+~\autocite{cock2009biopython}. Sequences were
multiply aligned using \software{MUSCLE} version 3.8.31
\autocite{edgar2004muscle}, and alignments were manually inspected with
\software{Seaview} version 4.4.2 \autocite{gouy2010seaview}. Phylogenies were
constructed from the nucleotide alignments by approximate maximum likelihood
using \software{FastTree2} version 2.1.7 \autocite{price2010fasttree} with the
\gls{GTR} model~\autocite{tavare1986some}. Transmission trees were estimated by
rooting and time-scaling the phylogenies by root-to-tip regression, using a
modified version of Path-O-Gen (distributed as part of
BEAST~\autocite{drummond2007beast}) as described
previously~\autocite{poon2015phylodynamic}.

### 2016-04-25 doc/thesis/preamble.tex
Small changes to objective
Finished results! Yes!
More writing.
Wrote discussion mostly
More writing
Panicked 1AM writing
Finished draft intro!

\documentclass[12pt]{report}
\usepackage[osf]{garamondx}
\usepackage[garamondx,cmbraces]{newtxmath}
\usepackage{marginnote}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nonumberlist]{glossaries}
\usepackage{glossary-mcols}
\usepackage{cleveref}
\usepackage{array}
\usepackage{textcomp}
\usepackage{float}

@@

\newcommand{\set}[1]{\left\lbrace#1\right\rbrace}
\newcommand{\sett}[1]{\{#1\}}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\vec}[1]{\mathbf{#1}}
\renewcommand{\star}[1]{#1^*}
\renewcommand{\d}{\mathrm{d}\,}

\newcommand{\N}{\mathcal{N}}
\renewcommand{\L}{\mathcal{L}}

@@

\DeclareMathOperator{\Exponential}{Exponential}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\Var}{Var}

\DeclareMathOperator{\inc}{in}
\DeclareMathOperator{\out}{out}

@@

\newacronym{MCMC}{MCMC}{Markov chain Monte Carlo}
\newacronym{ER}{ER}{Erd\H{o}s-R\'enyi}
\newacronym{BA}{BA}{Barab\'asi-Albert}
\newacronym{WS}{WS}{Watts-Strogatz}
\newacronym{SI}{SI}{susceptible-infected}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{PCA}{PCA}{principal components analysis}
\newacronym{SMC}{SMC}{sequential Monte Carlo}
\newacronym{ABC}{ABC}{approximate Bayesian computation}
\newacronym{nltt}{nLTT}{normalized lineages-through-time}
\newacronym{ltt}{LTT}{lineages-through-time}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{rv}{rv}{random variable}
\newacronym{ML}{ML}{maximum likelihood}
\newacronym{MAP}{MAP}{maximum \textit{a posteriori}}
\newacronym{GSL}{GSL}{GNU scientific library}
\newacronym{SIR}{SIR}{susceptible-infected-recovered}
\newacronym{IS}{IS}{importance sampling}
\newacronym{SIS}{SIS}{sequential importance sampling}
\newacronym{HMM}{HMM}{hidden Markov model}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{HPD}{HPD}{highest posterior density}
\newacronym{HIV}{HIV}{human immunodeficiency virus}
\newacronym{SARS}{SARS}{severe acute respiratory syndrome}
\newacronym{ODE}{ODE}{ordinary differential equation}
\newacronym{GLM}{GLM}{generalized linear model}

@@

>>>>

\documentclass[11pt]{report}
\usepackage[osf]{garamondx}
\usepackage[garamondx,cmbraces]{newtxmath}
\usepackage[top=1in, bottom=1in, inner=1.25in, outer=1in]{geometry}
\usepackage{setspace}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nonumberlist]{glossaries}
\usepackage{glossary-mcols}
\usepackage{cleveref}
\usepackage{array}
\usepackage{textcomp}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{fancyhdr}
\usepackage[dvipsnames]{xcolor}

@@

\newcommand{\set}[1]{\left\lbrace#1\right\rbrace}
\newcommand{\sett}[1]{\{#1\}}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\vec}[1]{\mathbf{#1}}
\renewcommand{\star}[1]{#1^*}
\renewcommand{\d}{\mathrm{d}\,}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\maps}{:}

\newcommand{\N}{\mathcal{N}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
\renewcommand{\L}{\mathcal{L}}

@@

\DeclareMathOperator{\Exponential}{Exponential}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\inc}{in}
\DeclareMathOperator{\out}{out}
\DeclareMathOperator{\ESS}{ESS}
\DeclareMathOperator{\nc}{nc}

@@

\newacronym{MCMC}{MCMC}{Markov chain Monte Carlo}
\newacronym{ER}{ER}{Erd\H{o}s-R\'enyi}
\newacronym{BA}{BA}{Barab\'asi-Albert}
\newacronym{WS}{WS}{Watts-Strogatz}
\newacronym{SI}{SI}{susceptible-infected}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{SVR}{SVR}{support vector regression}
\newacronym{kSVM}{kSVM}{kernel support vector machine}
\newacronym{kSVR}{kSVR}{kernel support vector regression}
\newacronym{PCA}{PCA}{principal components analysis}
\newacronym{kPCA}{kPCA}{kernel principal components analysis}
\newacronym{SMC}{SMC}{sequential Monte Carlo}
\newacronym{ABC}{ABC}{approximate Bayesian computation}
\newacronym{nltt}{nLTT}{normalized lineages-through-time}
\newacronym{ltt}{LTT}{lineages-through-time}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{rv}{rv}{random variable}
\newacronym{ML}{ML}{maximum likelihood}
\newacronym{MAP}{MAP}{maximum \textit{a posteriori}}
\newacronym{GSL}{GSL}{GNU scientific library}
\newacronym{SIR}{SIR}{susceptible-infected-recovered}
\newacronym{IS}{IS}{importance sampling}
\newacronym{SIS}{SIS}{sequential importance sampling}
\newacronym{HMM}{HMM}{hidden Markov model}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{HPD}{HPD}{highest posterior density}
\newacronym{HDI}{HDI}{highest density interval}
\newacronym{HIV}{HIV}{human immunodeficiency virus}
\newacronym{SARS}{SARS}{severe acute respiratory syndrome}
\newacronym{ODE}{ODE}{ordinary differential equation}
\newacronym{GLM}{GLM}{generalized linear model}
\newacronym{tasp}{TasP}{treatment as prevention}
\newacronym{MH}{MH}{Metropolis-Hastings}
\newacronym{GTR}{GTR}{generalized time-reversible}
\newacronym{ESS}{ESS}{expected sample size}
\newacronym{MSM}{MSM}{men who have sex with men}
\newacronym{ERGM}{ERGM}{exponential random graph model}

@@

\pagestyle{fancyplain}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\rfoot{\thepage}

\renewcommand\UrlFont{\color{Blue}\rmfamily}
